---
title: "BurglaryToPointOfInterestDistance"
date: "September 9, 2015"
output: html_document
---

```{r load-crimedata,echo=FALSE,warning=FALSE,message=FALSE,cache=TRUE}
library(sp)
library(rgeos)
library(rgdal)

setwd("/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/ExploratoryAnalysis/Features/")
# source("importCrimeData.R")
# filePath <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/ExploratoryAnalysis/CPD_DWH/"
# fileName <- "X_BURGLARY_POINTS_08_14.csv"
# BurglaryData <- importCrimeData(filePath,fileName)
# row.names(BurglaryData) <- NULL
# 
# shapefilePath.new <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/CPDShapeFiles/new/"
# beat_new.rg <- readOGR(paste0(shapefilePath.new,"cpd_beats"), "cpd_beats")
# district_new.rg <- readOGR(paste0(shapefilePath.new, "cpd_districts"),"cpd_districts")
# # centroids 
# Crd.beat <- coordinates(beat_new.rg)
# Crd.district <- coordinates(district_new.rg)
# 
# ## Match old and new beat 
# # Matching the old beat records and the new one by finding which new beat polygon the point falls in
# # and then re-assign that beat number to that crime record. 
# # Therefore all crime beat records have a unified reference which is the new beat map.
# source("DataPolyMatching.R")
# Match1 <- DataMatching2(BurglaryData,beat_new.rg,area="BEAT")
# BurglaryData <- Match1$CrimeData
# Match2 <- DataMatching2(BurglaryData,district_new.rg,area="DISTRICT")
# BurglaryData <- Match2$CrimeData
# 
# ## Aggregated by "beat" and add 'holiday' attribute
# source("HolidayChart.R")
# BurglaryData$HOLIDAY <- sapply(BurglaryData$DATEOCC,FUN=holidays)
# BurglaryData$HOLIDAY <- factor(BurglaryData$HOLIDAY)

load("MatchedBurglaryData_portal.RData")
```

We take a subset of burglary data so that contains only 2008-2014 records to relieve computational burden. The regular grid is made by setting up a 120 $\times$ 160 grid then constraint it by keeping those inside the city boundary, in which 7724 grids are included.

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=6, fig.height=4.5, cache=TRUE}
source("POIFunction.R")
Path.GIS <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/GISData/"
YearList <- 2008:2014
BurglaryData.sub <- subset(BurglaryData,YEAR>=2008&YEAR<=2014)

Path.city <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/GISData/City_Boundary"
city.shp <- readOGR(Path.city,"City_Boundary") 
ncell.x <- 120
ncell.y <- 160
X_range <- city.shp@bbox[1,]
Y_range <- city.shp@bbox[2,]
grd.full <- expand.grid(list(X_COORD=seq(X_range[1],X_range[2],length.out=ncell.x),
                             Y_COORD=seq(Y_range[1],Y_range[2],length.out=ncell.y)))
coordinates(grd.full) = ~X_COORD+Y_COORD # convert to SpatialPoints
proj4string(grd.full) <- proj4string(city.shp)

grdInCity <- over(grd.full,as(city.shp,"SpatialPolygons"))
grd.inCity <- grd.full[!is.na(grdInCity)]

RegGrd <- as.data.frame(grd.inCity@coords)

plot(grd.inCity,pch=16,cex=0.1,main="Regular grids within city boundary",cex.main=0.8)
box(which = "plot", lty = "solid")
```

Now we explore burglary in proximity to some spatial points of interest. They are all calculated in the same fasion that for each incident/grid we calculate the Euclidean distance between that point to the closest landmarks.

```{r cpd-station, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.cpd_station <- paste0(Path.GIS,"police_stations_poly")
cpd_station.shp <- readOGR(Path.cpd_station,"police_stations")
plot(cpd_station.shp, main="CPD stations")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,cpd_station.shp,"Dist2CPDstation",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,cpd_station.shp,"Dist2CPDstation")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2CPDstation,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest police stations")
hist(RegGrd$Dist2CPDstation,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest police stations")
```

```{r school, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.school <- paste0(Path.GIS,"School_Grounds")
school.shp <- readOGR(Path.school,"School_Grounds")
plot(school.shp, main="Schools")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,school.shp,"Dist2School",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,school.shp,"Dist2School")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2School,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest schools")
hist(RegGrd$Dist2School,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest schools")
```

```{r park, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.park <- paste0(Path.GIS,"Parks_Aug2012")
park.shp <- readOGR(Path.park,"Parks_Aug2012")
plot(park.shp, main="Parks")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,park.shp,"Dist2Park",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,park.shp,"Dist2Park")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2Park,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest parks")
hist(RegGrd$Dist2Park,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest parks")
```

```{r hospital, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.hospital <- paste0(Path.GIS,"Hospitals")
hospital.shp <- readOGR(Path.hospital,"Hospitals")
plot(park.shp, main="Hospitals")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,hospital.shp,"Dist2Hospital",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,hospital.shp,"Dist2Hospital")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2Hospital,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest hospitals")
hist(RegGrd$Dist2Hospital,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest hospitals")
```

```{r library, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.library <- paste0(Path.GIS,"Libraries")
library.shp <- readOGR(Path.library,"Libraries")
plot(park.shp, main="Libraries")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,library.shp,"Dist2Library",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,library.shp,"Dist2Library")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2Library,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest libraries")
hist(RegGrd$Dist2Library,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest libraries")
```

```{r cta-stop, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.CTA_stop <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/GISData/CTA_BusStops"
CTA_stop.shp <- readOGR(Path.CTA_stop,"CTA_BusStops")
prj <- paste("+proj=tmerc +lat_0=36.66666666666666 +lon_0=-88.33333333333333 +k=0.9999749999999999",
              "+x_0=300000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs +ellps=GRS80 +towgs84=0,0,0")
CTA_stop.shp <- spTransform(CTA_stop.shp, CRSobj=prj)

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,CTA_stop.shp,"Dist2BusStop",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,CTA_stop.shp,"Dist2BusStop")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2BusStop,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest bus stops")
hist(RegGrd$Dist2BusStop,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest bus stops")
```

```{r cta-route, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.CTA_route <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/GISData/CTA_Routes"
CTA_route.shp <- readOGR(Path.CTA_route,"CTA_Routes")
plot(CTA_route.shp, main="CTA routes")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,CTA_route.shp,"Dist2CTAroute",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,CTA_route.shp,"Dist2CTAroute")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2CTAroute,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest CTA routes")
hist(RegGrd$Dist2CTAroute,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest CTA routes")
```

```{r cta-rail, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.CTA_rail <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/GISData/CTA_RailLines"
CTA_rail.shp <- readOGR(Path.CTA_rail,"CTA_RailLines")
plot(CTA_rail.shp, main="CTA rails")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,CTA_rail.shp,"Dist2CTArail",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,CTA_rail.shp,"Dist2CTArail")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2CTArail,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest CTA rails")
hist(RegGrd$Dist2CTArail,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest CTA rails")
```

```{r street, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=10, fig.height=4.5, cache=TRUE}
Path.street <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/GISData//Major_Streets"
street.shp <- readOGR(Path.street,"Major_Streets")
plot(street.shp, main="Major Streets")
plot(city.shp,add=TRUE)
box(which = "plot", lty = "solid")

BurglaryData.sub <- Crime2POIdist(BurglaryData.sub,street.shp,"Dist2Street",YearList=YearList,MonthList=1:12)
RegGrd <- Grd2POIdist(RegGrd,street.shp,"Dist2Street")

par(mfrow=c(1,2))
hist(BurglaryData.sub$Dist2Street,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between burglary incidents and nearest major streets")
hist(RegGrd$Dist2Street,breaks=100,xlab="Distance (ft)", freq=FALSE,cex.main=0.6,
     main="Histogram of distances between regular grids and nearest major streets")
```